Incentivizing Exploration by Heterogeneous Users

### 2018-02-13 algorithm.tex
minor edits

The algorithm achieving the claimed bounds of
Theorem~\ref{thm:main-overview} is fairly simple.
It mostly allows agents to exploits, but when an arm is estimated to
be extremely unlikely to be pulled by any agents,
the algorithm incentivizes such an arm with a payment that is so high
that the next agent is guaranteed to pull the arm.
This way, the algorithm ensures that each arm is pulled at least a
certain minimum number of times.

More precisely, the algorithm divides time into \emph{phases}
$s = 1, 2, 3, \ldots$.
Phase $s$ ends when each arm has been pulled at least $s$ times.
An arm $i$ is \emph{payment-eligible} at time $t$ (in phase $s$)
if both of the following hold:

\begin{itemize}
\item $i$ has been pulled at most%
\footnote{in fact: exactly, since the algorithm entered phase $s$}
$s-1$ times up to time $t$, i.e., $\NumPull{t}{s} \leq s-1$.
\item Based on the current estimates \ArmEV{t}{i'} of all arms'
attribute vectors, the probability of pulling arm $i$ is less than $1/\log(s)$. \bccomment{Do we need to mention what if $\frac{1}{\log(s)}>1$? }
Let $\PullProb{t}{i} = \Prob[\AgV \sim \AgentDist]{\AgV \cdot \ArmEV{t}{i} > \AgV
  \cdot \ArmEV{t}{i'} \mbox{ for all } i' \neq i}$
be the probability that arm $i$ will be pulled
by the next (random) agent based on the current estimates. 
Then, the second criterion for payment-eligibility is that
$\PullProb{t}{i} < 1/\log(s)$.
\end{itemize}

@@

The high-level idea of the proof of Theorem~\ref{thm:main-overview}
is the following.
Because the algorithm ensures that each arm is pulled
``frequently enough,''
the estimates \ArmEV{t}{i} become exponentially more accurate in the
phase number $s$.
Thus, the fraction of agents who misidentify their best arm decreases
exponentially as well.
Because for each arm, enough agents have a true preference for this
arm, once the arms' attribute vectors are learned well enough,
the algorithm will not need to incentivize any arm pulls any more,
resulting in a payment bound independent of $T$.
Similarly, the regret will decrease exponentially, and mostly accrue
due to agents close to the border making mistakes in their arm choices.
The detailed analysis following this outline is somewhat subtle, though,
due to the dependencies between the agents' arm pulls and the
estimates which in turn are based on past arm pulls.
We begin with several technical lemmas that are used for both the
payment and regret bounds.

To capture the probability masses of problematic and non-problematic
agents, we define the following:
$\AlmostTied{\delta} = \Prob[\AgV \sim \AgentDist]{%
\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq \delta}$
is the total probability mass (across all arms) of agents
whose (actual) preference for their best arm over their second-best
arm is at most $\delta$.
The minimum probability mass of users who prefer their best arm
sufficiently strongly over their second-best arm is denoted by
$\ClearPref{\delta} = \min_{i}
\Prob[\AgV \sim \AgentDist]{\Best{\AgV}=i,%
\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) > \delta}$.
Notice that while \AlmostTied{\delta} (being a measure of problematic
agents) is defined as a sum over all arms,
while  \ClearPref{\delta} (being a measure of non-problematic agents)
is defined as a minimum over arms.
Also notice that $\ClearPref{0} = \MinProb$.
\dkcomment{We should stress somewhere that we assume that actual ties
  have measure 0.}

@@

\begin{proof}
Let $S$ be the set of arms that have been pulled at most
$s-1$ times at the start of phase $s$.
Thus, phase $s$ ends when each $i \in S$ has been pulled at least once.
Consider any time step $t$.
If $S$ contains at least one payment-eligible arm $i$,
then some payment-eligible arm will be pulled with probability 1,
and every payment-eligible arm is in $S$.
In particular, the size of $S$ decreases with probability 1.
If $S$ contains no payment-eligible arm,
then by definition of payment-eligibility,
each arm $i \in S$ has probability at least $1/\log(s)$ of being pulled.
In particular, the size of $S$ decreases with probability at least $1/\log(s)$.

>>>>

The algorithm achieving the claimed bounds of
Theorem~\ref{thm:main-overview} is fairly simple.
It mostly allows agents to exploit, but when an arm is \pfdelete{estimated to
be extremely} \pfedit{sufficiently} unlikely to be pulled\pfdelete{by any agents},
the algorithm incentivizes this arm with a payment high enough
to guarantee the next agent pulls it.
This way the algorithm ensures each arm is pulled often enough.

More precisely, the algorithm divides time into \emph{phases}
$s = 1, 2, 3, \ldots$.
Phase $s$ \pfcomment{starts?} ends when each arm has been pulled at least $s$ times.
We indicate the start time of phase $s$ by $t_s$.
\pfcomment{According to the notation we had before, in the .tex file below here,
phase $s$ starts when I have $s$ pulls of each arm.  Also, we should make sure things are clear regarding the free pull at time $0$.}
An arm $i$ is \emph{payment-eligible} at time $t$ (in phase $s$)
if both of the following hold:
\pfdelete{
We call time $t_{s}=\min_{t}\{\forall i, \NumPull{t}{i} \geq s\}$ the
\emph{starting point of the $s^{th}$ phase}.
We call the set of times $[t_{s}, t_{s+1})$ the
\emph{$n^{th}$ phase}.
}

\begin{itemize}
\item $i$ has been pulled at most%
\footnote{in fact: exactly, since the algorithm entered phase $s$}
$s-1$ times up to time $t$, i.e., $\NumPull{t}{s} \leq s-1$.
\item Based on the current estimates \ArmEV{t}{i'} of all arms'
attribute vectors, the probability of pulling arm $i$ is less than $1/\log(s)$. \bccomment{Do we need to mention what if $\frac{1}{\log(s)}>1$? }
Let $\PullProb{t}{i} = \Prob[\AgV \sim \AgentDist]{\AgV \cdot \ArmEV{t}{i} > \AgV
  \cdot \ArmEV{t}{i'} \mbox{ for all } i' \neq i}$
be the probability that arm $i$ will be pulled
by the next (random) agent based on the current estimates. 
Then, the second criterion for payment-eligibility is that
$\PullProb{t}{i} < 1/\log(s)$.
\pfedit{$1/\log(1)$ is treated as $\infty$, and if $1/\log(s) \ge 1$ then this condition is automatically met.}
\end{itemize}

@@

The high-level idea of the proof of Theorem~\ref{thm:main-overview}
is the following.
Because the algorithm ensures that each arm is pulled
``frequently enough,''
the estimates \ArmEV{t}{i} become exponentially more accurate in the
phase number $s$.
Thus, the fraction of agents who misidentify their best arm decreases
exponentially as well.
Because for each arm, enough agents have a true preference for this
arm, once the arms' attribute vectors are learned well enough,
the algorithm will not need to incentivize any more,
resulting in a payment bound independent of $T$.
Similarly, the regret will decrease exponentially, and mostly accrue
due to agents close to the border \pfcomment{we could clarify what border} making mistakes in their arm choices.
The detailed analysis following this outline is somewhat subtle, though,
due to the dependencies between the agents' arm pulls and the
estimates which in turn are based on past arm pulls.
We begin with several technical lemmas that are used for both the
payment and regret bounds.

To capture the probability masses of problematic and non-problematic
agents \pfcomment{in another pass we could seek to clarify what we mean by being problematic in the context of the border statement above}, we define the following:
$\AlmostTied{\delta} = \Prob[\AgV \sim \AgentDist]{%
\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq \delta}$
is the total probability mass (across all arms) of agents
whose (actual) preference for their best arm over their second-best
arm is at most $\delta$.
\pfcomment{We can check through the whole paper on whether we have $\AgV \sim \AgentDist$ in the subscript or just $\AgV \sim \AgentDist$.  I think we have only $\AgentDist$ in the problem description.}
The minimum probability mass of users who prefer their best arm
sufficiently strongly over their second-best arm is denoted by
$\ClearPref{\delta} = \min_{i}
\Prob[\AgV \sim \AgentDist]{\Best{\AgV}=i,%
\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) > \delta}$.
Notice that \AlmostTied{\delta} (being a measure of problematic
agents) is defined as a sum over all arms,
while  \ClearPref{\delta} (being a measure of non-problematic agents)
is defined as a minimum over arms.
Also notice that $\ClearPref{0} = \MinProb$.
\dkcomment{We should stress somewhere that we assume that actual ties
  have measure 0.}

@@

\begin{proof}
Let $S$ be the set of arms that have been pulled at most
$s-1$ times at the start of phase $s$.
Thus, phase $s$ \pfedit{has ended} when each $i \in S$ has been pulled at least once.
Consider any time step $t$.
If $S$ contains at least one payment-eligible arm $i$,
then some payment-eligible arm will be pulled with probability 1,
and every payment-eligible arm is in $S$.
In particular, the size of $S$ decreases with probability 1.
If $S$ contains no payment-eligible arm,
then by definition of payment-eligibility,
each arm $i \in S$ has probability at least $1/\log(s)$ of being pulled.
In particular, the size of $S$ decreases with probability at least $1/\log(s)$.
